Foundations of ecological and evolutionary change

### 2025-06-09 foundations_of_ecoevo.Rmd
Some more minor updates right before making final edits for Victor

---
title: "Foundations of ecological and evolutionary change"
author: |
  | A. Bradley Duthie$^{1,a,*}$ and Victor J. Luque$^{2,a}$
date: '[1] Department of Biological and Environmental Sciences, University of Stirling, Stirling, Scotland [2] Department of Philosophy, University of Valencia, Valencia, Spain [*] Corresponding author: alexander.duthie@stir.ac.uk [a] Equal contribution'
output:
  word_document: default
  pdf_document:
      keep_tex: yes
  html_document:
    df_print: paged
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lineno}
- \linenumbers
- \modulolinenumbers[2]
linestretch: 2
bibliography: refs.bib
link-citations: yes
linkcolor: blue
csl: the-american-naturalist.csl
biblio-style: apalike
---

@@

For example, if mean fitness at $t$ is 2 (i.e., $\bar{w} = 2$), then half as many individuals will have contributed to $\Omega$ in $t + 1$ than would have if $\bar{w} = 1$ (i.e., there are $N$ individuals at $t$ and $2N$ individuals at $t + 1$).
We therefore need to multiply the mean trait value $\bar{z}'$ (at $t + 1$) by the mean fitness $\bar{w}$ (at $t$) to recover the mean contribution of the $N$ individuals at $t$ to the total $\Omega$.
Consequently,

@@

Our fundamental equation is complete and exact.
It therefore implicitly includes any effects of density dependence on population growth (see Box 2), or any social effects on evolutionary change (see Box 1).
Both of these effects can be made explicit by specifying how other individuals in a population affect birth and death of a focal individual.
Assuming that effects of individuals are additive, independent, and identical, we can derive well-established logistic growth equations and Hamilton's rule from eqn 1 (Supporting Information S1).

>>>>

---
title: "Foundations of ecological and evolutionary change"
author: |
  | A. Bradley Duthie$^{1,a,*}$ and Victor J. Luque$^{2,a}$
date: '[1] Department of Biological and Environmental Sciences, University of Stirling, Stirling, Scotland [2] Department of Philosophy, University of Valencia, Valencia, Spain [*] Corresponding author: alexander.duthie@stir.ac.uk [a] Equal contribution'
output:
  pdf_document:
      keep_tex: yes
  word_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lineno}
- \linenumbers
- \modulolinenumbers[2]
linestretch: 2
bibliography: refs.bib
link-citations: yes
linkcolor: blue
csl: the-american-naturalist.csl
biblio-style: apalike
---

@@

For example, if mean fitness at $t$ is 2 (i.e., $\bar{w} = 2$), then half as many individuals will have contributed to $\Omega$ in $t + 1$ than would have if $\bar{w} = 1$ (i.e., there are $N$ individuals at $t$ and $2N$ individuals at $t + 1$).
We therefore need to multiply the mean trait value $\bar{z}'$ (at $t + 1$) by the mean fitness $\bar{w}$ (at $t$) to recover the mean contribution of the $N$ individuals at $t$ to the total $\Omega$ [this is noted in the context of population genetics by @Ewens2014].
Consequently,

@@

Our fundamental equation is complete and exact.
It therefore implicitly includes any effects of density dependence on population growth (see Box 2), or any social effects on evolutionary change (see Box 1).
Both of these effects can be made explicit by specifying how other individuals in a population affect birth and death of a focal individual.
Assuming that effects of individuals are additive, independent, and identical, we can derive well-established logistic growth equations and multi-level selection from eqn 1 (Supporting Information S1).

### 2025-06-09 foundations_of_ecoevo.tex
Some more minor updates right before making final edits for Victor

For example, if mean fitness at \(t\) is 2 (i.e., \(\bar{w} = 2\)), then
half as many individuals will have contributed to \(\Omega\) in
\(t + 1\) than would have if \(\bar{w} = 1\) (i.e., there are \(N\)
individuals at \(t\) and \(2N\) individuals at \(t + 1\)). We therefore
need to multiply the mean trait value \(\bar{z}'\) (at \(t + 1\)) by the
mean fitness \(\bar{w}\) (at \(t\)) to recover the mean contribution of
the \(N\) individuals at \(t\) to the total \(\Omega\). Consequently,

@@

>>>>

For example, if mean fitness at \(t\) is 2 (i.e., \(\bar{w} = 2\)), then
half as many individuals will have contributed to \(\Omega\) in
\(t + 1\) than would have if \(\bar{w} = 1\) (i.e., there are \(N\)
individuals at \(t\) and \(2N\) individuals at \(t + 1\)). We therefore
need to multiply the mean trait value \(\bar{z}'\) (at \(t + 1\)) by the
mean fitness \(\bar{w}\) (at \(t\)) to recover the mean contribution of
the \(N\) individuals at \(t\) to the total \(\Omega\) (this is noted in
the context of population genetics by \citeproc{ref-Ewens2014}{Ewens
2014}). Consequently,

@@

\bibitem[\citeproctext]{ref-Ewens2014}
Ewens, W. J. 2014.
\href{https://doi.org/10.1007/s10539-013-9412-0}{Grafen, the price
equations, fitness maximization, optimisation and the fundamental
theorem of natural selection}. Biology and Philosophy 29:197--205.
